-------------------------------------------------------------

#Activity monitoring
##author: Dominika Ra

date: 11 kwietnia 2022

output: html_document


-------------------------------------------------------------

R version:
4.1.1 (2021-08-10) -- "Kick Things"
Copyright (C) 2021 The R Foundation for Statistical Computing
Platform: x86_64-w64-mingw32/x64 (64-bit)

-------------------------------------------------------------
##1. Code for reading in the dataset and/or processing the data


```{r setup, include= FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r 1}
data<-read.csv("activity.csv")
names(data)
head(data)
str(data)
summary(data)
```
##I create a datebase without NA's
```{r 0}
data1<-data[which(data$step>=0), ]
dim(data1)
```

##2. Histogram of the total number of steps taken each day


```{r 2}
plot(tapply(data1$steps,data1$date,sum),type ="l",xlab="Day of the analysis",ylab = "Sum of steps in a day", main = "The total number of steps taken each day")
```


##3. Mean and median number of steps taken each day

```{r 3}
library(dplyr)
by_date<-group_by(data1,date)
summarize(by_date,mean(steps),median(steps))
```




##4. Time series plot of the average number of steps taken

```{r 4}
plot(tapply(data1$steps,data1$date,mean),type ="l",xlab="Day of the analysis",ylab = "Mean of steps in a day", main = "The mean value of steps taken each day")
```



##5. The 5-minute interval that, on average, contains the maximum number of steps

```{r 5}
plot(tapply(data1$steps,data1$interval,mean),type ="l",xlab="Interval number",ylab = "Mean of steps in interval", main = "The mean number of steps taken in interval")
plot(tapply(data1$steps,data1$interval,max),type ="l",xlab="Interval number",ylab = "Maximum value for steps in interval", main = "The max number of steps taken in interval")
```



##6. Code to describe and show a strategy for imputing missing data

```{r bledy}
sum(is.na(data$steps))
```

```{r total}
sum(is.na(data$steps))/nrow(data)*100
```

###The total number of rows with NA = ` r bledy`. NA's account for ` r total` % of all rows.

###The NA's were fill in with mean for a interval
```{r na}
data1<-data[which(data$step>=0), ]
by_interval<-group_by(data1,interval)
inter_df<-data.frame(summarize(by_interval,mean(steps)))
head(inter_df)
data_na<-data[is.na(data$steps), ]
data_na_ok<-merge(data_na,inter_df,by.x = "interval",by.y = "interval")
data_na_ok1<-select(data_na_ok, mean.steps.,date,interval)
data_na_ok2<-rename(data_na_ok1,steps = mean.steps.)
data_total<-rbind(data1,data_na_ok2)
dim(data_total)
head(data_total)
```
##7. Histogram of the total number of steps taken each day after missing values are imputed
```{r 7}
plot(tapply(data_total$steps,data_total$date,sum),type ="l",xlab="Day of analysis",ylab = "Sum of steps in a day", main = "The total number of steps taken each day")
```



##8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r 8}
data_total_w<-mutate(data_total,date=as.Date(date))
data_total_w1<-mutate(data_total_w,weekdays(date))
slownik<-data.frame(weekdays=c("poniedziałek","wtorek","środa","czwartek","piątek","sobota","niedziela"),rodzaj = c("week","week","week","week","week","weekend","weekend"))
slownik
data_total_w2<-merge(data_total_w1,slownik,by.x="weekdays(date)",by.y="weekdays")
data_total_w3<-select(data_total_w2,steps,interval,rodzaj)
by_interval_total<-group_by(data_total_w3,rodzaj,interval)
a<-summarize(by_interval_total,mean(steps))
a<-rename(a,msteps = "mean(steps)")
library(ggplot2)
qplot(interval,msteps,data=a,facets=rodzaj~.,geom = c("line") ,main = "Average number of steps taken per 5-minute interval across weekdays and weekends") 
```



